---
title: "Simulation Lecture Coding Example"
author: "Sarah Teichman"
date: "5/3/2022"
output: html_document
---

Note: this is edited to be slightly different from the example in class in order to be more instructive. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Study the the variance of the sample mean for draws from a binomial distribution with sample sizes $n=10$ and $n=1000$. Consider the binomial distribution with $5$ trials and probabilities $p$ in the set $\{0.2, 0.4, 0.6, 0.8\}$. 

```{r}
# set nsim 
nsim <- 10000 
# write function to draw data and calculate variance once 
draw_data_once <- function(n, p) {
  draw <- rbinom(n = n, size = 5, prob = p) 
  return(mean(draw))
}
# create sample sizes vector 
n_vec <- c(10, 1000) 
# create probabilities vector 
p_vec <- seq(0.2, 0.8, 0.2)
# create matrix to hold results 
results_10 <- matrix(NA, nrow = nsim, ncol = 4)
results_1000 <- matrix(NA, nrow = nsim, ncol = 4)
# run simulation 
for (i in 1:nsim) {
  for (prob in 1:length(p_vec)) {
    results_10[i, prob] <- draw_data_once(10, p_vec[prob])
    results_1000[i, prob] <- draw_data_once(1000, p_vec[prob])
  }
}
# calculate results 
# make vectors to hold variances of sample means across trials 
var_of_means10 <- rep(NA, length(p_vec))
var_of_means1000 <- rep(NA, length(p_vec))
# loop over each probability of success
for (i in 1:length(p_vec)) {
  # compute variance of sample means for one n and p combination
  var_of_means10[i] <- var(results_10[, i])
  var_of_means1000[i] <- var(results_1000[, i])
}
# plot results 
# make dataframe to put data to plot inside 
plot_df <- data.frame(mean_var = c(var_of_means10,
                                   var_of_means1000),
                      sample_size = c(rep(10, 4), rep(1000, 4)),
                      probability = rep(p_vec, 2))
# recode sample size as a factor to use as facet labels 
plot_df$sample_size <- as.factor(plot_df$sample_size) %>%
  fct_recode("Sample Size: 10" = "10", 
             "Sample Size: 1000" = "1000")
# make faceted line plot 
# here I chose to facet instead of using color because the 
# two sets of variances are on such different scales 
# by adding "scales = 'free'" to "facet_wrap()" I can have 
# different y axis scales for my facets 
ggplot(plot_df, aes(y = mean_var, x = probability,)) +
  geom_line() + 
  facet_wrap(~ as.factor(sample_size), scales = "free") + 
  labs(x = "Probability", 
       y = "Variance of Sample Mean",
       title = "Variance of Sample Mean by n and p") + 
  theme(plot.title = element_text(hjust = 0.5))
```

From this plot, we can see that the variance of the sample mean is a function of both the probability of success and the sample size. The variance of the sample mean is largest for probabilities near $1/2$ and smaller near probabilities close to $0$ and $1$. The variance of the sample mean is smaller for larger sample sizes, and an increase in sample size of $100$ times leads to a $100$ times decrease in the variance of the sample mean. 